---
title: "Untitled"
author: "wilfried Dossou-Yovo"
date: "2022-10-10"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```


# Sensitivity of CO2 emission rate to temperature in cranberry soils 
*Wilfried Dossou-Yovo, Serge Étienne Parent,Noura Ziadi, Bertrand Normand, and Léon Étienne parent*

```{r}
print(Sys.Date())
```

## 1. Objective

This notebook generate the result of CO2 data analysis. Data set contains a collection of soil caracteristics, measured co2 emission collected from incubation study. Soil samples was collected from two cranberry fied stand of eastern canada. Incubation study was carried out at Agriculture and Agri-food Canada(sainte-foy, quebec,qc) from February to Mai 2019. The aim of this study was to measure CO2 emission rates in cranberry soils of Eastern Canada as related to soil temperature and depth

## 2. Statistical questions

In addition to data exploration, this notebook will answer the following statistical questions.
 
1. What is the influence of soil depth and temperature on CO2 emission?
2. Can Arrhenius equation and Q10 be useful to describe temperature sensitivity of carbon decomposition across layers? 

## 3. Packages

We need package [tidyverse](https://www.tidyverse.org/) which loads a set of packages for easy data manipulation(Ex: [dplyr](https://www.rdocumentation.org/packages/dplyr/versions/0.7.8)) and visualization (ex: [ggplot2](https://www.rdocumentation.org/packages/ggplot2/versions/3.2.1)). We also use [ggpubr](https://www.rdocumentation.org/packages/ggpubr/versions/0.1.1) to customise publication ready plot, [ggpmisc](https://www.rdocumentation.org/packages/ggpmisc/versions/0.2.17) and [grid](https://www.rdocumentation.org/packages/grid/versions/3.6.2) are useful packages as extensions to [ggplot2](https://www.rdocumentation.org/packages/ggplot2/versions/3.2.1).

```{r}
#install.packages(c("ggpubr", "tidyverse", "tidymodels", "plotly"))
library(tidyverse)
library(tidymodels)
library(ggpubr)
library(plotly)

```

### 4. Import data
We load two data `data_pot` and `data_co2` involved in our anylisis. `data_pot` contained details about sites sampling, soil sampling(soil depth, weight, water content and bulk density), laboratory incubation temperature while `data_co2` contained details about laboratory incubation time, co2 emission and jar masson details.
`data_co2` was combined with `data_pot` with `left_join` function

```{r}
data_pots <- read_csv2('data/pots.csv')
data_co2 <- read_csv('data/co2.csv')
data_co2 <- data_co2 |>
left_join(data_pots, by = "ID pot")
data_pots
```

## 5. Some calculations 

Several variables have been added to our data in order to proceed for analysis. The added variables are the following: `Temperature (Kelvin)`, `Molar Volume (L/mol)`, `Headspace Volume (mL)`, `Dry soil weight (g)`, `CO2 emission (ug/h/g)`, `CO2 emission (mg/kg)`, `decomposition rate K`, `lnK`and `1/T(T = Temperature(Kelvin)`

```{r}
container_volume <- 250 # mL
gas_constant <- 8.31446 # unit
atm_pressure_n <- 101.325

data_co2 <- data_co2 %>%
  mutate(`Temperature (K)` = `Temperature (°C)` + 273,
         `Total carbon (Mg/ha)` = `Bulk density (g/cm3)` * `Carbone(%)` * 10,
         `Molar Volume (L/mol)` = gas_constant * `Temperature (K)` / atm_pressure_n,
         `Headspace Volume (mL)` = container_volume - (`Soil weight (g)` / `Bulk density (g/cm3)`), # 250 mL is the volume of the container
         `Dry soil weight (g)` = `Soil weight (g)` - (`Soil weight (g)` * `Water content (%)` / 100),
         `CO2 emission (ug/h/g)` = (`Final CO2 (ppm)` - `Initial CO2 (ppm)`) * 0.000001 * 44000000 /
           `Molar Volume (L/mol)` * (`Headspace Volume (mL)` / 1000) * (12 / 44) /
           `Time final (h)` / `Dry soil weight (g)`,
         `CO2 emission (mg/kg)` = `CO2 emission (ug/h/g)` * 24 * `Time (days)`,
         `CO2 emission (Mg/ha)` = `CO2 emission (mg/kg)` * `Bulk density (g/cm3)` * 10 * 0.0001,
         K = log(`Total carbon (Mg/ha)` / (`Total carbon (Mg/ha)` - `CO2 emission (Mg/ha)`)) / `Time (days)`,
         lnK = log(K),
         `1/T` = 1 / `Temperature (K)`)

#write_xlsx(data_co2, "data/data_co2_leon.xlsx")
```

## 6. Exploratory data analysis

### 6.1 Histogram

```{r}
New.labs <- c("10°C", "20°C", "30°C") # Change labels 
names(New.labs) <- c("10", "20", "30")

New.labs_b <-  c("[0-10 cm]", "[10-20 cm]", "[20-30 cm]") # Change labels
names(New.labs_b) <- c("10", "20", "30")
```

```{r}
library(plotly)
ggplotly(  
  data_co2 |>
    ggplot() +
    geom_histogram(aes(x = log10(`CO2 emission (ug/h/g)`)), bins = 150)
)
```

Data contains some outliers, let remove them

```{r}
data_co2_clean <- data_co2 |> 
  mutate(log_tr = log10(`CO2 emission (ug/h/g)`)) |>
  filter(log_tr > -3.06 & log_tr < -0.33) |>
  drop_na()
```

Now data look well distributed
```{r}
ggplotly(  
  data_co2_clean |>
    ggplot() +
    geom_histogram(aes(x = log10(`CO2 emission (ug/h/g)`)), bins = 100) 
)
```


### 6.2 Correlations

```{r}
data_co2_clean |>
  select(`Time (days)`, `Depth (cm)`, `Temperature (°C)`, `CO2 emission (ug/h/g)`) |>
  corrr::correlate() |>
  corrr::focus(`CO2 emission (ug/h/g)`) |>
  mutate(term = fct_reorder(term, `CO2 emission (ug/h/g)`)) |>
  ggplot(aes(x = `CO2 emission (ug/h/g)`, y= term)) +
  geom_col(width = 0.2) +
  labs(x = bquote(~CO[2]~ 'emision ('*mu~'g'~ h^-1~g^-1*')')) +
  theme_bw()
```

### 6.3 Boxplot

```{r}
options(repr.plot.width = 6, repr.plot.height = 7)
pg <- ggplot(data=data_co2_clean, aes(x = `Time (days)`,y = `CO2 emission (ug/h/g)` )) +
  geom_boxplot(aes(group = factor(`Time (days)`))) + 
  facet_grid(`Depth (cm)` ~ `Temperature (°C)`,   scales = "free", 
             labeller = labeller(`Depth (cm)` = New.labs_b, `Temperature (°C)` = New.labs))+ 
  labs(x = "Incubation time (days)", y = bquote(~CO[2]~ 'emision ('*mu~'g'~ h^-1~g^-1*')'))
pg 
ggsave("figures/Boxplot.png", width = 6, height = 7, dpi = 600)# export plot high resolution
```

## 7.What is the influence of soil depth and temperature on CO2 emission?

### 7.1 Build model: linear regression

```{r}
model_rec <-  data_co2_clean |>
  recipe(`CO2 emission (ug/h/g)` ~ ., data_co2) |> 
  step_select(`CO2 emission (ug/h/g)`, `Time (days)`, Sites, 
              `Depth (cm)`, `Temperature (°C)`) |>
  step_log(all_outcomes(), base = 10) |>
  step_dummy(Sites) |>
  step_normalize(all_numeric(), -all_outcomes() ) |>
  prep()

data_co2_preprocessed <-  juice(model_rec)
model_spec <- linear_reg() |>
  set_engine("lm")
```

#### Fit model
```{r}
model_fit <- model_spec |>
  fit(`CO2 emission (ug/h/g)` ~ ., data_co2_preprocessed) 
```


#### Exploring model results
```{r}
tidy(model_fit)
```

```{r}
glance(model_fit)
```

#### Inspect the model
```{r}
par(mfrow=c(2,2)) # plot all 4 plots in one

plot(model_fit$fit, 
     pch = 16,    
     col = '#006EA1')
```

#### Prediction
```{r}
prediction <-  model_fit |>
  predict(data_co2_preprocessed)
```

#### collect Metrics
```{r}
rmse <-  data_co2_preprocessed |>
  bind_cols(prediction) |>
  rmse(`CO2 emission (ug/h/g)`, .pred)
rmse
rmse <-  round(as.numeric(rmse[1,3]), 2)
```

```{r}
rsq <-  data_co2_preprocessed |>
  bind_cols(prediction) |>
  rsq(`CO2 emission (ug/h/g)`, .pred)
rsq
rsq <- round(as.numeric(rsq[1,3]), 2)
```

```{r}
options(repr.plot.width = 4, repr.plot.height = 2)
px <- data_co2_preprocessed |>
  bind_cols(prediction) |>
  ggplot(aes(x = `CO2 emission (ug/h/g)`, y = .pred)) +
  geom_point(size = 1, alpha = .3) +
  geom_label(aes(x = -3, y = -.6),
             vjust = 1, hjust = 0, size = 2, label.size = 0.1,
             label = paste("R² =",  rsq, "\nRMSE =", rmse)) +
  geom_abline() +
  theme_pubr()  +
  theme(axis.title=element_text(size=7),
        axis.line = element_line(size = 0.1),
        axis.ticks = element_line(size = 0.1),
        axis.text = element_text(size = 6)) +
  labs(x= bquote("Observed log"~CO[2]~ 'emission ('*mu~'g'~ h^-1~g^-1*')'~"rate") , y = bquote("Predicted log"~CO[2]~ 'emission ('*mu~'g'~ h^-1~g^-1*')'~"rate"))
px  
ggsave("figures/Observed and predicted co2 emission.png", width = 4, height = 2.2, dpi = 600)
```

### 7.3 Variable coefficient and confidence intervals

```{r}
options(repr.plot.width = 8, repr.plot.height = 2)
term_rename <- tibble(term = c("`Time (days)`", "`Depth (cm)`", "`Temperature (°C)`", "Sites_PF45"),
                      name_corrected = c("Time (d)", "Layer (cm)", "Temperature (°C)", "Sites"))

h <- broom::tidy(model_fit, conf.int = TRUE) |>
  dplyr::filter(term != "(Intercept)") |>
  left_join(term_rename, by = "term") |>
  mutate(term_rename = fct_reorder(name_corrected, estimate)) |>
  ggplot(aes(estimate, term_rename)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.1,
                 size=0.5) +
  labs(x = "Coefficient", y = "") +
  theme_light() +
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=14)) # Time (d)
h
ggsave("figures/Linear-model-Co2_with_site.png", width = 8, height = 2, dpi = 600)
```

### 7.3 Prediction model of CO2 emission in cranberry soils in three-layer positions (0-10 cm, 10-20 cm, 20-30 cm) and at three temperatures (10, 20 and 30oC).

```{r}
New.labs <- c("10°C", "20°C", "30°C") # Change labels 
names(New.labs) <- c("10", "20", "30")

New.labs_b <-  c("[0-10 cm]", "[10-20 cm]", "[20-30 cm]") # Change labels
names(New.labs_b) <- c("10", "20", "30")

options(repr.plot.width = 8, repr.plot.height = 6)
pl <- data_co2_clean |>
  bind_cols(10^prediction) |>
  ggplot(aes(x = `Time (days)`, y = `CO2 emission (ug/h/g)`)) +
  geom_point(size = 2, alpha = .3) +
  facet_grid(`Depth (cm)` ~ `Temperature (°C)`, scales = "free", labeller = labeller(`Depth (cm)` = New.labs_b, `Temperature (°C)` = New.labs)) +
  geom_line(aes(x = `Time (days)`, y = `.pred`)) +
  scale_y_log10() +
  theme_bw() +
  theme(strip.text = element_text(size = 12), axis.text=element_text(size=12),
        axis.title=element_text(size=14),
        axis.title.y = element_text(size=14)) +
  xlab("Incubation time (d)") + ylab(bquote(~CO[2]~ 'emission rate ('*mu~'g'~ h^-1~g^-1*')'))
pl
ggsave("figures/CO2 emission.png", plot= pl, width = 7, height = 5, dpi = 600)
```

## 8. What is the temperature sensitivity across cranberry soil layers?

### 8.1 Fit of Arrhenius equation

The Arrhenius equation has been used to describe temperature sensitivity to CO2 emission. The Arrhenius equation was computed as follows: 



$$k = A e^{{\frac{-Ea}{RT}}}$$

$$log \left( k \right) = log \left( A e^{\frac{-Ea}{RT}} \right)$$

$$log \left( k \right) = log \left( A \right) + log \left(e^{\frac{-Ea}{RT}} \right)$$

$$log \left( k \right) = log \left( A \right) - \frac{1}{T} \times \left(\frac{Ea}{R}\right)$$

Where $A$ is the pre-exponential factor and $Ea$ is activation energy assumed to be independent of temperature, $R$ is the universal gas constant and $T$ is absolute temperature (Kelvin)


```{r}
models_co2 <- data_co2 %>%
  group_by(`Depth (cm)`) %>%
  summarise(linmod = list(lm(lnK ~ `1/T`)))
models_co2
```

```{r}
linmod_coef <- list()
for (i in seq_along(models_co2$linmod)) linmod_coef[[i]] <- models_co2$linmod[[i]]$coefficients
linmod_coef <- do.call(rbind.data.frame, linmod_coef)
names(linmod_coef) <- c("Intercept", "Slope")
linmod_coef <- bind_cols(unique(data_co2["Depth (cm)"]), linmod_coef)
linmod_coef
```


```{r}
options(repr.plot.width = 12, repr.plot.height = 6)
plot_co2 <- data_co2_clean %>%
  ggplot(aes(x = `1/T`, y = lnK)) +
  facet_grid(~`Depth (cm)`, labeller = labeller(`Depth (cm)` = New.labs_b)) +
  geom_boxplot(aes(group = factor(`1/T`)), outlier.shape = NA) +
  stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~")), label.x = 0.00331, label.y = -7) +
  geom_abline(data = linmod_coef, aes(intercept = Intercept, slope = Slope), lwd = 1) +
  labs(y = bquote("log(k)" ~d^-1)) +
  theme_bw() +
  theme(strip.text = element_text(size = 14), axis.text=element_text(size=14),
        axis.title=element_text(size=14)) 
plot_co2
ggsave("figures/Arrhénius équation.png", plot = plot_co2, width = 8, height = 4, dpi = 600)# export plot high resolution
```


### 8.2 Activation Energy computation

```{r}
Activation_energy <- tibble(
  Soil_layers = c("10", "20", "30"),
  intercept = NA,
  slope = NA,
  adj_r_sq = NA
)
```

```{r}
lm_arrhenius <- for (i in 1:nrow(Activation_energy)) {
  
  lm_Activation_energy <- data_co2_clean %>%
    filter(`Depth (cm)` == Activation_energy$Soil_layers[i]) %>%
    lm(lnK ~ `1/T`, data = .)
  
  # intercept
  Activation_energy$intercept[i] <- coef(lm_Activation_energy)[1]
  
  # Slope
  Activation_energy$slope[i] <- coef(lm_Activation_energy)[2]
  
  # statistics
  Activation_energy$adj_r_sq[i] <- summary(lm_Activation_energy)$adj.r.squared
}
```

```{r}
R = 8.3144621 / 1000 # Gas constant Kj/mol/K 
Activation_energy <-  Activation_energy %>%
  mutate(Ea = -slope * R) %>%
  select(Soil_layers, adj_r_sq, Ea)
Activation_energy
```

### 8.3 Computing K median in order to compute Q10 value accross soil depth
```{r}
K_median <- aggregate(K ~ `Sites` + `Time (days)` + `Depth (cm)` + `Temperature (°C)`, data = data_co2_clean, FUN = median)

K_median <- K_median %>%
  pivot_wider(names_from = `Temperature (°C)`, values_from = K)


K_median$Q_20_10 <- K_median$`20` / K_median$`10`
K_median$Q_30_20 <- K_median$`30` / K_median$`20`

K_median <- K_median %>%
  na.omit(K_median)
data_Q10 <- gather(data = K_median, key = `Temperature range`, value = Q10, c(`Q_20_10`, `Q_30_20`),
                   factor_key=TRUE)

stat_Q10 <- data_Q10 |>
  group_by(`Depth (cm)`) |>
  get_summary_stats(Q10)
stat_Q10
```

```{r}
New.labs_c <- c("[10°C - 20°C]", "[20°C - 30°C]") # Change labels 
names(New.labs_c) <- c("Q_20_10", "Q_30_20")

options(repr.plot.width = 8, repr.plot.height = 4)
data_Q10 |> 
  mutate(`Layers` = as.character(`Depth (cm)`)) |>
  ggplot(aes(x = `Time (days)`, y = `Q10`)) +
  facet_grid(`Temperature range`~`Depth (cm)`, 
             labeller = labeller(`Depth (cm)` = New.labs_b,
                                 `Temperature range` = New.labs_c)) +
  geom_smooth(method = "lm", se = TRUE, color = "Black") +
  geom_point(size = 1.5, alpha = 0.6) +
  labs(x = "Time(d)", y = bquote(Q[10])) +
  theme_bw() + 
  theme(strip.text = element_text(size = 14), axis.text=element_text(size=14),
        axis.title=element_text(size=14)) 
ggsave("figures/Variation of Q10 across layers.png", width = 8, height = 4, dpi = 600)# export plot high resolution
```

## 9. Soil description

### 9.1 Soil layers properties

Import data

```{r}
data_carbon_credit <- read_csv2('data/data_carbon_credit.csv')
data_carbon_credit <- data_carbon_credit |>
  mutate(`C:N ratio` = `Carbone (%)` / `Nitrogen (%)`)
```

Some calculations
```{r}
mean_sd_CoverN <- data_carbon_credit |>
  group_by(`Layer (cm)`) |>
  summarize(mean_C_over_N = mean(`C:N ratio`, na.rm = TRUE), 
                        se_C_over_N = sd(`C:N ratio`, na.rm = TRUE)/sqrt(length(!is.na(`C:N ratio`))))
mean_sd_CoverN

data_carbon_credit |> get_summary_stats(`C stock (kg m-3)`)

data_carbon_credit |> 
  group_by(`Layer (cm)`) |>
get_summary_stats(`C stock (kg m-3)`)

data_carbon_credit |> 
  group_by(`Layer (cm)`) |>
get_summary_stats(`C:N ratio`)
```

```{r}

plot_desc <- function(y, ylab){
  New.labs_c <-  c("Site/A9", "Site/45") # Change labels
names(New.labs_c) <- c("Belanger/ A9", "Fortier/ 45")
  ggplot(data_carbon_credit, aes(`Layer (cm)`, y)) +
    geom_boxplot() +
facet_grid( . ~ `Location`, scales = "free", labeller = labeller(`Location` = New.labs_c)) +
    theme_bw() +
theme(strip.text = element_text(size = 11), axis.text=element_text(size=11),
        axis.title=element_text(size=11)) +
    labs(y = ylab)
  }

plot1 <- plot_desc(data_carbon_credit$`C stock (kg m-3)`, bquote("C stock (kg" ~m^-3~")"))
plot2 <- plot_desc(data_carbon_credit$`C:N ratio`, "C:N ratio")
plot3 <- plot_desc(data_carbon_credit$`Bulk density (kg m-3)`, # m-3
                   "Bulk density (kg"~m^-3~")")
plot4 <- plot_desc(data_carbon_credit$pHCaCl2, bquote(pHCaCl[2]))
plot5 <- plot_desc(data_carbon_credit$`Total porosity`, "Total porosity")
plot6 <- plot_desc(data_carbon_credit$`Water content (%)`, "Water content (%)")

options(repr.plot.width = 8, repr.plot.height = 6)
figure <- ggarrange(plot1, plot2, plot3, plot4, plot5, plot6,
                    labels = c("A", "B", "C", "D", "E", "F"), label.x = 0.05, label.y = 1.01,
                    ncol = 2, nrow = 3)
figure
ggsave("figures/Soil description.png", width = 8, height = 5, dpi = 600)# export plot high resolution
```

## 9.2 C:N ratio in alternate sublayers of sand and organic matter

Data loading

```{r}
Carbon_credit <- read_csv2('data/data_carbon_sublayer.csv')
Carbon_credit
```

C:N ratio computation

```{r}
Carbon_credit <- Carbon_credit %>%
mutate(`C/N` = C_pourc/N_pourc)
```

Generating the plots

```{r}
options(repr.plot.width=8, repr.plot.height=4)
New.labs_d <-  c("Site/A9", "Site/45") # Change labels
names(New.labs_d) <- c("Belanger/A9", "Fortier/45")

ggplot(data=Carbon_credit, aes(x= `Depht (cm)`, y= `C/N`)) +
  facet_grid(.~Site, labeller = labeller(`Site` = New.labs_d)) +
  geom_line(linetype = "twodash") +
  geom_point(aes(shape = `Soil texture`, fill = `Soil texture`), size = 3) +
  scale_shape_manual(values=c(21, 21))+
  scale_fill_manual(values = c("#000000", "#FFFFFF")) +
  scale_y_continuous(breaks = 5*0:1000,
                     expand = expand_scale(add = 5)) +
  scale_x_continuous(breaks = 5*0:1000,
                     expand = expand_scale(add = 5)) +
  theme(strip.text = element_text(face = "bold"), axis.text=element_text(face = "bold"),
        axis.title=element_text(face = "bold") , legend.title= element_text(face = "bold"),
        legend.text = element_text(face = "bold")) +
  labs(y= "C/N Ratio") +
  coord_flip()
ggsave("figures/(C_over_N).png", width = 8, height = 4, dpi = 800)
```

